Equalizer 2.0 --\\Convergence of a Parallel Rendering Framework

### 2018-01-17 paperIEEE.tex
Redone introduction.
git-svn-id: https://dev.ifi.uzh.ch/vmml/svn/Equalizer16@89 db7c848b-7755-4df2-b763-8796797afa52

In this systems paper, we present an overview of the integration of research and
commercial developments into the Equalizer parallel rendering framework first
presented in \cite{EMP:09}. While individual research using the system has been
presented since, many important features have not previously been published and
are presented here. In particular, the integration of parallel rendering
research, APIs for developing complex distributed applications, and many
individual features make Equalizer a unique, open source framework to develop
visualization applications for any type of setup and use case.

We do present novel algorithms for parallel rendering which did not warrant a
separate publication: pixel and subpixel decompositions, dynamic frame
resolution, tunable sort-first load-balancing parameters, frame rate
equalization, thread synchronization modes for multi-threaded rendering, a
powerful abstraction for multi-view rendering on arbitrary display setups,
dynamic focus distance and asymmetric eye positions for Virtual Reality,
parallel pixel streaming to tiled display walls, as well as a fully-fledged data
distribution API with compression and reliable multicast.

The rest of this paper is structured as follows: First we introduce new related
work since the introduction of \textsf{Equalizer}. The main body of this paper
presents new performance features, virtual reality algorithms, usability
features to build complex applications, main novel features of the underlying
\textsf{Collage} network library and a quick overview of the main
\textsf{Equalizer}-based applications. A result section presents new experiments
not previously published, followed by the discussion and conclusion.

>>>>

The Equalizer parallel rendering framework as first presented in~\cite{EMP:09} has
demonstrated its general versatility and the usefulness of its minimally-intrusive 
programming design in a variety of applications and projects.
In particular, the integration of large-scale parallel rendering algorithms,
APIs for developing complex distributed applications, and many
individual features make Equalizer a unique, open source framework to develop
visualization applications for virtually any type of setup and use case.
While individual applications and case studies using the framework, as well as
new algorithms and system components extending it have been presented
since the initial release of Equalizer, many important features and functionalities
have not previously been published and are presented here.
In this report, we thus present an updated comprehensive review of the
integration of research, application use cases and commercial developments
with respect to Equalizer as well as novel comparative experimental results
of its scalability features.

In particular, we present novel algorithms for parallel rendering which did not yet
appear in a separate publication, including pixel and sub-pixel decompositions,
dynamic frame resolution, tunable sort-first load-balancing parameters, frame-rate
equalization, thread synchronization modes for multi-threaded rendering, a
powerful abstraction for multi-view rendering on arbitrary display setups,
dynamic focus distance and asymmetric eye positions for VR,
parallel pixel streaming to tiled display walls, as well as a fully-fledged data
distribution API with compression and reliable multicast.

The remainder of this paper is structured as follows: First we provide an update
on related work since the introduction of \textsf{Equalizer}. The main body of this paper
then presents new performance features, VR algorithms, usability
features to build complex applications, main novel features of the underlying
\textsf{Collage} network library and a quick overview of the main
\textsf{Equalizer}-based applications. A result section presents new experiments
not previously published, followed by the discussion and conclusion.
